---
title: "R Notebook"
output: html_notebook
---

In this notebook, the baseline model for the data set called "data" will be documented.

First, the data will be arranged by date and then divided into a training data set (01.07.2013 until 31.07.2017) and a validation data set (01.08.2017 until 31.07.2018).

```{r}
# import libraries
library("dplyr")

# load csv-file with all data
data <- read_csv("0_DataPreparation/data.csv")

# sort data by Datum
data_sorted <- data %>% arrange(Datum)
head(data_sorted)

# divide data into training and validation set
training_data <- data_sorted %>% filter(Datum <= "2017-07-31")
validation_data <- data_sorted %>% filter(Datum >= "2017-08-01")

# check new data sets
head(training_data)
head(validation_data)
```

Before the linear regression, all variables will be looked at in case they need to be adjusted:

- Warengruppe -> use as a category
- KielerWoche -> too many missing values and therefore currently not useful 
- Bewoelkung -> is linear, use as is
- Temperatur -> is not linear, needs to be used as a binned category
- Windgeschwindigkeit -> is linear, use as is
- Wettercode -> too many missing values and therefore currently not useful
- Schulferien -> too many missing values and therefore currently not useful
- Wochenende -> boolean, use as is

```{r}
# How to bin the temperature data?
# looking at the min, max and mean
summary(training_data$Temperatur)  # data ranges from -8.475 to 31.438 degrees

# show distribution of data
hist(training_data$Temperatur)

# approach one: create 8 bins of equal width for the temperature
training_data1 <- training_data %>%
  mutate(Temperatur_bins = cut(Temperatur, 8, include.lowest = TRUE))

# check new added temperature_bins column and its unique values
head(training_data1)
training_data1 %>% distinct(Temperatur_bins) %>% arrange(Temperatur_bins)

```

The last step will be setting up the linear model equation with the training data set and searching for the model with the best adjusted R-squared value.

```{r}
# model 1
mod1 <- lm(Umsatz ~ as.factor(Warengruppe) + Temperatur_bins, training_data1)
summary(mod1)
```

Model 1: Adjusted R-squared:  0.6876

```{r}
# model 2
mod2 <- lm(Umsatz ~ as.factor(Warengruppe) + Temperatur_bins + Bewoelkung, training_data1)
summary(mod2)
```

Model 2: Adjusted R-squared:  0.6877

The R-squared is just slightly better than model 1, but the p-value of Bewoelkung indicates that it is not as relevant to the model as most of the other variables. It will be not be used in the next model.

```{r}
# model 3
mod3 <- lm(Umsatz ~ as.factor(Warengruppe) + Temperatur_bins + Windgeschwindigkeit, training_data1)
summary(mod3)
```

Model 3: Adjusted R-squared:  0.6878

The R-squared is again just slightly better than model 1, the p-value of Windgeschwindigkeit indicates that it is relevant to the model, but less relevant than most other variables.

```{r}
# model 4
mod4 <- lm(Umsatz ~ as.factor(Warengruppe) + Temperatur_bins + Windgeschwindigkeit + Wochenende, training_data1)
summary(mod4)
```

Model 4: Adjusted R-squared:  0.713 

This is so far the best R-squared. The p-value of the variable Windgeschwindigkeit indicates, that it is less relevant and will not be used in the next model.


```{r}
# model 5
mod5 <- lm(Umsatz ~ as.factor(Warengruppe) + Temperatur_bins + Wochenende, training_data1)
summary(mod5)
```

Model 5: Adjusted R-squared:  0.7129

This models R-squared differs from model 4 by only 0.0001, but works without the variable Windgeschwindigkeit. With one variable less but almost the same R-squared it can be considered the best model for now.
