---
title: "Data Characteristics"
output: html_notebook
---

This notebook will provide a short look on the different variables of the dataset and it's main characteristics.

```{r}
# import libraries
library(readr)
library(lubridate)
library(ggplot2)
library(dplyr)
```

```{r}
# load data
data <- read_csv("../0_DataPreparation/data.csv")
head(data)
tail(data)
```

```{r}
# check for missing values
missing_values <- sapply(data, function(x) sum(is.na(x)))
print(missing_values)
```

Creating a bar plot that shows the average sales (Umsatz) across all product groups by school holidays (Schulferien).

```{r}
# group the data by Schulferien (0 or 1) and calculate the average sale
mean_umsatz_schulferien <- data %>%
  group_by(Schulferien) %>%
  summarise(Durchschnittsumsatz = mean(Umsatz, na.rm = TRUE))

# show grouped dataset
print(mean_umsatz_schulferien)

# change values of Schulferien for better readability
mean_umsatz_schulferien$Schulferien[mean_umsatz_schulferien$Schulferien == 0] <- "Nein"
mean_umsatz_schulferien$Schulferien[mean_umsatz_schulferien$Schulferien == 1] <- "Ja"

# create bar plot
bar_schulferien <- ggplot(mean_umsatz_schulferien) +
  aes(x=Schulferien, y=Durchschnittsumsatz) +
  geom_bar(stat="identity", color="darkblue", fill="darkblue") +
  geom_text(aes(label=round(Durchschnittsumsatz, digits = 2)), vjust=1.6, color="white", size=4.5)+
  ggtitle("Durchschnittlicher Umsatz während der offiziellen Schulferien") +
    theme_minimal()
bar_schulferien

# save plot as image file
ggsave("bar_schulferien.jpeg", plot = bar_schulferien)
```

Creating a bar plot that shows the average sales (Umsatz) across all product groups by weekend (Wochenende).

```{r}
# group the data by Wochenende (0 or 1) and calculate the average sale
mean_umsatz_wochenende <- data %>%
  group_by(Wochenende) %>%
  summarise(Durchschnittsumsatz = mean(Umsatz, na.rm = TRUE))

# show grouped dataset
print(mean_umsatz_wochenende)

# change values of Wochenende for better readability
mean_umsatz_wochenende$Wochenende[mean_umsatz_wochenende$Wochenende == 0] <- "Nein"
mean_umsatz_wochenende$Wochenende[mean_umsatz_wochenende$Wochenende == 1] <- "Ja"

# create bar plot
bar_wochenende <- ggplot(mean_umsatz_wochenende) +
  aes(x=Wochenende, y=Durchschnittsumsatz) +
  geom_bar(stat="identity", color="darkred", fill="darkred") +
  geom_text(aes(label=round(Durchschnittsumsatz, digits = 2)), vjust=1.6, color="white", size=4.5)+
  ggtitle("Durchschnittlicher Umsatz am Wochenende") +
    theme_minimal()
bar_wochenende

# save plot as image file
ggsave("bar_wochenende.jpeg", plot = bar_wochenende)
```

Creating a bar plot that shows the average sales (Umsatz) across all product groups by weekday (Wochentag).

```{r}
# check type of date
class(data$Datum)

# Datum is a character/string and needs to be changed to date
data$Datum <- as.Date(data$Datum, format ="%Y-%m-%d")

# Creating a new variable for weekday (Wochentag)
data$Wochentag <- weekdays(data$Datum)
head(data)

# Categorize Wochentage with as.factor to ensure the correct order of the days of the week
data$Wochentag <- factor(data$Wochentag, levels = c("Montag", "Dienstag", "Mittwoch", "Donnerstag", "Freitag", "Samstag", "Sonntag"))

# group the data by Wochentag and calculate the average sale
mean_umsatz_wochentag <- data %>%
  group_by(Wochentag) %>%
  summarise(Durchschnittsumsatz = mean(Umsatz, na.rm = TRUE))

# show grouped dataset
print(mean_umsatz_wochentag)

# create bar plot
bar_wochentag <- ggplot(mean_umsatz_wochentag) +
  aes(x=Wochentag, y=Durchschnittsumsatz) +
  geom_bar(stat="identity", color="darkgreen", fill="darkgreen") +
  geom_text(aes(label=round(Durchschnittsumsatz, digits = 2)), vjust=1.6, color="white", size=3.5)+
  ggtitle("Durchschnittlicher Umsatz je Wochentag") +
    theme_minimal()
bar_wochentag

# save plot as image file
ggsave("bar_wochentag.jpeg", plot = bar_wochentag)
```

Add error bars for Umsatz to the bar plot.

```{r}

# Calculate confidence interval (ic) (from r graph gallery)
  # Calculates mean, sd, se and iC
my_sum <- data %>%
  group_by(Wochentag) %>%
  summarise( 
    n=n(),
    mean=mean(Umsatz, na.rm = TRUE),
    sd=sd(Umsatz, na.rm = TRUE)
  ) %>%
  mutate( se=sd/sqrt(n))  %>%
  mutate( ic=se * qt((1-0.05)/2 + .5, n-1))
   # check
print(my_sum)

# create bar plot with error bars
bar_wochentag_errorbar <- ggplot(my_sum) +
  aes(x=Wochentag, y=mean) +
  geom_bar(stat="identity", color="darkgreen", fill="darkgreen") +
  geom_errorbar(aes(x=Wochentag, ymin=mean-ic, ymax=mean+ic), width=0.4, colour="orange", alpha=0.9, linewidth=1.5) +
  ggtitle("Durchschnittlicher Umsatz je Wochentag mit Konfidenzintervall") +
    theme_minimal()
bar_wochentag_errorbar

# save plot as image file
ggsave("bar_wochentag_errorbar.jpeg", plot = bar_wochentag_errorbar)
```

Create a bar plot with sales by weekday, separated by prodcut group (Warengruppe).

```{r}
# group by Wochentag + Warengruppe and calculate average Umsatz
mean_umsatz_wochentag_warengr <- data %>%
  group_by(Warengruppe, Wochentag) %>%
  summarise(Durchschnittsumsatz = mean(Umsatz, na.rm = TRUE))

# show grouped data
print(mean_umsatz_wochentag_warengr)

# create readable labels for Warengruppe
warengruppen_liste <- c(
  "1" = "Brot",
  "2" = "Brötchen",
  "3" = "Croissant",
  "4" = "Konditorei",
  "5" = "Kuchen",
  "6" = "Saisonbrot")

# check for missing values
sum(is.na(mean_umsatz_wochentag_warengr))

# delete rows with missing values
mean_umsatz_wochentag_warengr <- mean_umsatz_wochentag_warengr[
  complete.cases(mean_umsatz_wochentag_warengr), ]

# create bar plot
bar_wochentag_warengruppe <- ggplot(mean_umsatz_wochentag_warengr) +
  aes(x=Wochentag, y=Durchschnittsumsatz) +
  geom_bar(stat="identity", color="darkgreen", fill="darkgreen") +
  geom_text(aes(label=round(Durchschnittsumsatz, digits = 1)), vjust=1.2, color="white", size=2.5)+
  ggtitle("Durchschnittlicher Umsatz pro Warengruppe je Wochentag") +
  facet_wrap(vars(Warengruppe), nrow = 3, labeller = labeller(Warengruppe = warengruppen_liste)) +
    theme_minimal()
bar_wochentag_warengruppe

# save plot as image file
ggsave("bar_wochentag_warengruppe.jpeg", plot = bar_wochentag_warengruppe)
```
